---
title: "Cell interaction analysis for mice fibroblasts and dentric cells"
author: "Jiangyan Yu (jiangyan.yu@uni-bonn.de)"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r global_options}
knitr::opts_chunk$set(warning=FALSE, messgae=FALSE, fig.path='Figs/', results = "hide",fig.width=8, fig.height=4)
```

# Libraries

```{r library, include=FALSE, warning=FALSE, message=FALSE}
rm(list=ls())
#CRAN packages
list.of.packages <- c("readr",
                      "cowplot",
                      "useful",
                      "stringr",
                      "umap",
                      "ggplot2",
                      "reshape2",
                      "dplyr",
                      "tidyr",
                      "Matrix.utils",
                      "VGAM",
                      "Seurat",
                      "magrittr",
                      "RColorBrewer",
                      "ggpubr",
                      "tibble"
)

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)>0) install.packages(new.packages)

#BioconductoR packages
list.of.bioc.packages <- c("tximport",
                           "DESeq2",
                           "slingshot",
                           "flowCore",
                           "biomaRt",
                           "clusterProfiler",
                           "org.Hs.eg.db",
                           "org.Mm.eg.db",
                           "GO.db",
                           "GSEABase",
                           "nichenetr",
                           "DOSE",
                           # "VISION",
                           "scater",
                           "CARNIVAL",
                           "ComplexHeatmap")
new.packages.bioc <- list.of.bioc.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
 
if(length(new.packages.bioc)>0)if (!requireNamespace("BiocManager")) install.packages("BiocManager")
BiocManager::install(new.packages.bioc, update = FALSE)

lapply(c(list.of.packages,list.of.bioc.packages), require, character.only = TRUE)

rm(list.of.packages,new.packages,list.of.bioc.packages,new.packages.bioc)
```

# working.dir and functions

```{r working directory}
working.dir = "/home/yu.j/sciebo/Projects/Zhang_DC_project_ownerJY/"

# 
function_dir = "/home/yu.j/sciebo/general_scripts/Global_ref_data/"
# published_data_dir = "/home/yu.j/sciebo/general_scripts/publised_datasets/"
# global_ref_dir = "/home/yu.j/sciebo/general_scripts/Global_ref_data/"
# gsea_pathway_dir = "/home/yu.j/sciebo/general_scripts/Global_ref_data/"
# 
# source(paste0(function_dir,"general_functions.R"))
```

## nichenetr wrapped function

download needed database from "zenodo.org/record/3260758":
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
ligand_tf_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_tf_matrix.rds"))
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
sig_network = readRDS(url("https://zenodo.org/record/3260758/files/signaling_network.rds"))
gr_network = readRDS(url("https://zenodo.org/record/3260758/files/gr_network.rds"))

```{r wrapped nichenetr function}
Zhang_DC_nichenetr_all_genes = function(seurat.object, receiver.cells, sender.cells,DE_table_receiver, label.receiver.cells, label.sender.cells, top_n_target_per_ligand, geneset_oi_avg_log2FC, best_ligand_top_n){
  RNA.counts=seurat.object@assays$RNA@counts
  receiver.cells=receiver.cells
  sender.cells=sender.cells
  DE_table_receiver=DE_table_receiver
  label.receiver.cells=label.receiver.cells
  label.sender.cells=label.sender.cells
  top_n_target_per_ligand=top_n_target_per_ligand
  geneset_oi_avg_log2FC=geneset_oi_avg_log2FC
  best_ligand_top_n=best_ligand_top_n
  
  ### import nichenet database
  # Nichenetr packages
  ligand_target_matrix = readRDS(paste(function_dir,"nichenetr_ligand_target_matrix.rds", sep = "/"))
  
  ##convert from human to mouse
  colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
  rownames(ligand_target_matrix) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()
  ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]
  
  lr_network = readRDS(paste(function_dir,"nichenetr_lr_network.rds",sep = "/"))
  
  lr_network$from_mouse <- lr_network$from %>% convert_human_to_mouse_symbols()
  lr_network$to_mouse <- lr_network$to %>% convert_human_to_mouse_symbols()
  
  weighted_networks = readRDS(paste(function_dir,"nichenetr_weighted_networks.rds",sep = "/"))
  weighted_networks_m = weighted_networks
  weighted_networks_m$lr_sig = weighted_networks_m$lr_sig %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% tidyr::drop_na()
  weighted_networks_m$gr = weighted_networks_m$gr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% tidyr::drop_na()
  
  weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))
  
  weighted_networks_lr = weighted_networks_lr %>% mutate(from_mouse = convert_human_to_mouse_symbols(from), to_mouse = convert_human_to_mouse_symbols(to)) %>% tidyr::drop_na()
  
  ligand_tf_matrix = readRDS(paste(function_dir,"nichenetr_ligand_tf_matrix.rds", sep = "/"))
  colnames(ligand_tf_matrix) = ligand_tf_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
  rownames(ligand_tf_matrix) = ligand_tf_matrix %>% rownames() %>% convert_human_to_mouse_symbols()
  ligand_tf_matrix = ligand_tf_matrix %>% .[!is.na(rownames(ligand_tf_matrix)), !is.na(colnames(ligand_tf_matrix))]
  
  sig_network = readRDS(paste(function_dir,"nichenetr_signaling_network.rds", sep = "/"))
  sig_network$from <- sig_network$from %>% convert_human_to_mouse_symbols()
  sig_network$to <- sig_network$to %>% convert_human_to_mouse_symbols()
  
  gr_network = readRDS(paste(function_dir,"nichenetr_gr_network.rds", sep = "/"))
  gr_network$from <- gr_network$from %>% convert_human_to_mouse_symbols()
  gr_network$to <- gr_network$to %>% convert_human_to_mouse_symbols()
  
  ### define genes of interests
  expressed_genes_receiver = RNA.counts %>% .[,receiver.cells] %>% apply(1,function(x){sum(x>0)/length(x)}) %>% .[. >=0.1] %>% names()
  
  receptors = lr_network %>% pull(to_mouse) %>% unique()
  expressed_receptors = intersect(receptors, expressed_genes_receiver)
  
  background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]
  
  ## define gene of interest
  # DE_table_receiver = FindMarkers(object = seurat.object, ident.1 = receiver.cells, min.pct = 0.10) %>% rownames_to_column("gene")
  
  geneset_oi = DE_table_receiver %>% dplyr::filter(p_val_adj <= 0.05 & avg_log2FC >= geneset_oi_avg_log2FC) %>% pull(gene) %>% .[. %in% rownames(ligand_target_matrix)] %>% as.character()
  
  ligands = lr_network %>% pull(from_mouse) %>% unique()
  
  sender_genes_sender = RNA.counts %>% .[,sender.cells] %>% apply(1,function(x){sum(x>0)/length(x)}) %>% .[. >=0.1] %>% names()
  expressed_ligands_sender = intersect(ligands,sender_genes_sender)
  expressed_ligands = expressed_ligands_sender
  
  potential_ligands = lr_network %>% dplyr::filter(from_mouse %in% expressed_ligands & to_mouse %in% expressed_receptors) %>% pull(from_mouse) %>% unique()
  
  ligand_activities = predict_ligand_activities(
    geneset = geneset_oi, 
    background_expressed_genes = background_expressed_genes, 
    ligand_target_matrix = ligand_target_matrix, 
    potential_ligands = potential_ligands)
  
  ligand_activities %>% arrange(-pearson) %>% mutate(rank = rank(plyr::desc(pearson)))
  ligand_activities
  
  best_upstream_ligands = ligand_activities %>% 
    top_n(best_ligand_top_n, pearson) %>%
    arrange(-pearson) %>% 
    pull(test_ligand) %>%
    unique()
  
  # DotPlot(seurat.object, features = best_upstream_ligands %>% rev(), cols = "Blue", split.by = "celltype") + RotatedAxis()
  
  # show histogram of ligand activity scores
  ## define cut-off for the number of selected ligands
  p_hist_lig_activity = ggplot(ligand_activities, aes(x=pearson)) + 
    geom_histogram(color="black", fill="darkorange")  + 
    # geom_density(alpha=.1, fill="orange") +
    geom_vline(aes(xintercept=min(ligand_activities %>% top_n(20, pearson) %>% pull(pearson))), color="red", linetype="dashed", size=1) + 
    labs(x="ligand activity (PCC)", y = "# ligands") +
    theme_classic()
  
  active_ligand_target_links_df = best_upstream_ligands %>% 
    lapply(get_weighted_ligand_target_links,geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = top_n_target_per_ligand) %>% 
    bind_rows() %>% drop_na()
  
  ### for visualisation
  active_ligand_target_links = prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.1) ## old cutoff was 0.25, new is 0.33
  
  order_ligands = intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% 
    rev() %>% make.names()
  order_targets = active_ligand_target_links_df$target %>% 
    unique() %>% 
    intersect(rownames(active_ligand_target_links)) %>% 
    make.names()
  rownames(active_ligand_target_links) = rownames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
  colnames(active_ligand_target_links) = colnames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
  
  vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% 
    t()
  
  p_ligand_target_network = vis_ligand_target %>% make_heatmap_ggplot(paste("Prioritized",label.sender.cells, "ligands"), paste("DE genes in", label.receiver.cells, "cells"), color = "purple",legend_position = "top", x_axis_position = "top",legend_title = "Regulatory potential") + scale_fill_gradient2(low = "whitesmoke",  high = "purple", breaks = c(0,0.005,0.01)) + theme(axis.text.x = element_text(face = "italic"))
  
  lr_network_top = lr_network %>% 
    dplyr::filter(from_mouse %in% best_upstream_ligands & to_mouse %in% expressed_receptors) %>% 
    distinct(from_mouse,to_mouse)
  
  best_upstream_receptors = lr_network_top %>% pull(to_mouse) %>% unique()
  
  lr_network_top_df_large = weighted_networks_lr %>% 
    dplyr::filter(from_mouse %in% best_upstream_ligands & to_mouse %in% best_upstream_receptors)
  
  lr_network_top_df = lr_network_top_df_large %>% tidyr::spread("from_mouse","weight",fill = 0)
  ###remove from and to columns due to mouse orgnism
  lr_network_top_df = lr_network_top_df[,c(-1,-2)]
  lr_network_top_matrix = lr_network_top_df %>% dplyr::select(-to_mouse) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to_mouse)
  
  # perform hierarchical clustering to order the ligands and receptors
  dist_receptors = dist(lr_network_top_matrix, method = "binary")
  hclust_receptors = hclust(dist_receptors, method = "ward.D2")
  order_receptors = hclust_receptors$labels[hclust_receptors$order]
  
  dist_ligands = dist(lr_network_top_matrix %>% t(), method = "binary")
  dist_ligands[is.na(dist_ligands)] = 0
  hclust_ligands = hclust(dist_ligands, method = "ward.D2")
  order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]
  
  order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix))
  order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix))
  
  vis_ligand_receptor_network = lr_network_top_matrix[order_receptors, order_ligands_receptor]
  rownames(vis_ligand_receptor_network) = order_receptors %>% make.names()
  colnames(vis_ligand_receptor_network) = order_ligands_receptor %>% make.names()
  
  p_ligand_receptor_network = vis_ligand_receptor_network %>% t() %>% make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", x_axis_position = "top",legend_title = "Prior interaction potential")
  
  print("Please mannually check the ligand_activities matrix to explore which ligands are biologically meaningful. Then set best_ligand_top_n accordingly.")
  
  return(list(p_hist_lig_activity=p_hist_lig_activity,
              p_ligand_target_network=p_ligand_target_network,
              p_ligand_receptor_network=p_ligand_receptor_network,
              order_receptors= order_receptors,
              order_ligands_receptor=order_ligands_receptor,
              ligand_activities = ligand_activities,
              active_ligand_target_links_df = active_ligand_target_links_df,
              lr_network_top_df_large = lr_network_top_df_large %>% rename(ligand = from_mouse, receptor = to_mouse)))  
}
```


# Raredon dataset

data were downloaded from https://advances.sciencemag.org/content/5/12/eaaw3851.abstract
following two samples are used: GSM3926539_mMouse, GSM3926539_mMouse

## load data and run Seurat

```{r seurat}
# mouse_data_dir = "/home/yu.j/sciebo/general_scripts/publised_datasets/GSE133747_mouse"
# mouse_ids = list.files(mouse_data_dir) # Should show barcodes.tsv, genes.tsv, and matrix.mtx
# 
# d10x.data <- sapply(mouse_ids, function(i){
#   mat <- readMM(paste(paste(mouse_data_dir,i, sep="/"),"raw.data.mtx.gz",sep="/"))
# 
# 
#   gene.names = read.delim(paste(paste(mouse_data_dir,i, sep="/"),"genes.tsv.gz",sep="/"),
#                            header = TRUE,
#                            stringsAsFactors = FALSE)
# 
#   #### input file with a header
#   barcode.names = read.delim(paste(paste(mouse_data_dir,i, sep="/"),"barcodes.tsv.gz",sep="/"),
#                            header = TRUE,
#                            stringsAsFactors = FALSE)
#   colnames(mat) = paste(i,barcode.names$x,sep="_")
#   rownames(mat) = gene.names$x
#   mat
# })
# 
# 
# experiment.data <- do.call("cbind", d10x.data)
# 
# lung_scRNA_mouse_Raredon <- CreateSeuratObject(
#   experiment.data,
#   project = "lung_scRNA_mouse_Raredon",
#   min.cells = 10)
#   # min.genes = 200,
#   # names.field = 2,
#   # names.delim = "\\-")
# 
# rm(experiment.data,d10x.data)
# 
# lung_scRNA_mouse_Raredon[["percent.mt"]] <- PercentageFeatureSet(lung_scRNA_mouse_Raredon, pattern = "^Mt")
# 
# VlnPlot(object = lung_scRNA_mouse_Raredon, features = c("nFeature_RNA","nCount_RNA","percent.mt"))
# 
# FeatureScatter(object = lung_scRNA_mouse_Raredon, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
# 
# FeatureScatter(object = lung_scRNA_mouse_Raredon, feature1 = "percent.mt", feature2 = "nFeature_RNA")
# 
# lung_scRNA_mouse_Raredon <- subset(lung_scRNA_mouse_Raredon, subset = nFeature_RNA > 500 & nFeature_RNA < 10000 & percent.mt < 1)
# 
# lung_scRNA_mouse_Raredon <- NormalizeData(lung_scRNA_mouse_Raredon, normalization.method = "LogNormalize", scale.factor = 10000)
# lung_scRNA_mouse_Raredon <- FindVariableFeatures(lung_scRNA_mouse_Raredon, selection.method = "vst", nfeatures = 2000)
# lung_scRNA_mouse_Raredon <- ScaleData(object = lung_scRNA_mouse_Raredon, features = rownames(x = lung_scRNA_mouse_Raredon))
# lung_scRNA_mouse_Raredon <- RunPCA(object = lung_scRNA_mouse_Raredon, features = VariableFeatures(object = lung_scRNA_mouse_Raredon),verbose = FALSE)
# lung_scRNA_mouse_Raredon <- JackStraw(object = lung_scRNA_mouse_Raredon, num.replicate = 100)
# lung_scRNA_mouse_Raredon <- ScoreJackStraw(object = lung_scRNA_mouse_Raredon, dims = 1:20)
# JackStrawPlot(object = lung_scRNA_mouse_Raredon, dims = 1:20)
# ElbowPlot(object = lung_scRNA_mouse_Raredon)
# 
# 
# lung_scRNA_mouse_Raredon <- FindNeighbors(object = lung_scRNA_mouse_Raredon, dims = 1:10)
# lung_scRNA_mouse_Raredon <- FindClusters(object = lung_scRNA_mouse_Raredon, resolution = 1.23)
# ### res1.23 gives 27 clusters
# lung_scRNA_mouse_Raredon <- RunUMAP(object = lung_scRNA_mouse_Raredon, dims = 1:10)
# DimPlot(object = lung_scRNA_mouse_Raredon, reduction = 'umap',label = FALSE)
# 
# lung_scRNA_mouse_Raredon.markers <- FindAllMarkers(lung_scRNA_mouse_Raredon, min.pct = 0.5, logfc.threshold = 0.4, only.pos = T)
# # plot.markers <- lung_scRNA_mouse_Raredon.markers[lung_scRNA_mouse_Raredon.markers$p_val_adj <0.001,]
# 
# test = lung_scRNA_mouse_Raredon@assays$RNA %>% .[rownames(.) %in% "Gpr183",] %>% t()  %>% as.data.frame()
# 
# test$group = 0
# 
# for(i in 1:nrow(test)){
#   tmp = as.numeric(test[i,1])
#   if(tmp==0){
#     test[i,2] = "0"
#   }else if(tmp >0 & tmp <=1){
#     test[i,2] = "0_1"
#   }else if(tmp >1 & tmp <=2){
#     test[i,2] = "1_2"
#   }else if(tmp >2 & tmp <=3){
#     test[i,2] = "2_3"
#   }
# }
# 
# lung_scRNA_mouse_Raredon = AddMetaData(lung_scRNA_mouse_Raredon,test$group,col.name = "Gpr183.exp")
# 
# rm(test)

# saveRDS(lung_scRNA_mouse_Raredon,file = paste0(working.dir,"lung_scRNA_mouse_Raredon_20210604.rds"))
```

# load saved seurat object

```{r load saved seurat object}
lung_scRNA_mouse_Raredon = readRDS(file = paste0(working.dir,"lung_scRNA_mouse_Raredon_20210604.rds"))
```

# nichenetr analysis

## define cDC2 cells

cluster 5

```{r define receiver cells}
DotPlot(lung_scRNA_mouse_Raredon, features = c("Itgae", "Ccl17", "Cd86","Mgl2","Gpr183"))
```

## define adventitial fibroblasts

cluster 9

```{r define sender cells}
DotPlot(lung_scRNA_mouse_Raredon, features = c("Pdgfra","Pi16","Col14a1", "Col1a1"))

FeaturePlot(lung_scRNA_mouse_Raredon, features = c("Pdgfra","Pi16","Col14a1", "Col1a1"), label = TRUE)
```

# nichenet analysis

```{r nichenetr analysis}
## receiver cells
Raredon.receiver.cells = subset(lung_scRNA_mouse_Raredon@meta.data,seurat_clusters==5 & Gpr183.exp!=0) %>% rownames()
## sender cells
Raredon.sender.cells = subset(lung_scRNA_mouse_Raredon@meta.data,seurat_clusters==9) %>% rownames()
## cell lables 
label.Raredon.sender.cells = "Col14a1+ fibroblast"
label.Raredon.receiver.cells = "Ccl17+Gpr183+ cDC2"
## DEGs
Raredon.DE_table_receiver = FindMarkers(object = lung_scRNA_mouse_Raredon, ident.1 = Raredon.receiver.cells, min.pct = 0.10,logfc.threshold = 0.25) %>% rownames_to_column("gene")

## 
## nichenetr analysis
Raredon.fibro_Gpr183cDC2_all = Zhang_DC_nichenetr_all_genes(lung_scRNA_mouse_Raredon, Raredon.receiver.cells, Raredon.sender.cells, Raredon.DE_table_receiver, label.Raredon.receiver.cells, label.Raredon.sender.cells,top_n_target_per_ligand = 100, geneset_oi_avg_log2FC=1, best_ligand_top_n=20)

Raredon.fibro_Gpr183cDC2_all$p_hist_lig_activity
Raredon.fibro_Gpr183cDC2_all$p_ligand_target_network
Raredon.fibro_Gpr183cDC2_all$p_ligand_receptor_network
```
# circos plot: nichenetr ligand-receptor 

## Calculate average ligand expression in sender cells

```{r prepare circos plot, warning=FALSE, message=FALSE, results='hide'}
## input objects
## 1. seurat object that you used for nichenetr analysis
## 2. Tsukui.fibro4_Gpr183cDC2_all: nichenet output including ligand_activities, active_ligand_target_links_df

# Calculate average ligand expression in sender cells 
top_ligands = Raredon.fibro_Gpr183cDC2_all$ligand_activities %>% 
    top_n(30, pearson) %>%
    arrange(-pearson) %>% 
    pull(test_ligand) %>%
    unique()

## remove a from the list
top_ligands = setdiff(top_ligands,"a")



### define sender specific ligands

ligand_type_indication_df = tibble(
  ligand_type = c(rep("Ligands", times = top_ligands %>% length())),
  ligand = top_ligands)



### Define the ligand-target links of interest

circos_setup = list()
modifed_df = Raredon.fibro_Gpr183cDC2_all$lr_network_top_df_large
colnames(modifed_df)[5] = "target"

active_ligand_target_links_df = modifed_df %>% mutate(target_type = "cDC2-DEGs") %>% inner_join(ligand_type_indication_df) # if you want to make circos plots for multiple gene sets, combine the different data frames and differentiate which receptor belongs to which gene set via the receptor type

cutoff_include_all_ligands = active_ligand_target_links_df$weight %>% quantile(0.20)

active_ligand_target_links_df_circos = active_ligand_target_links_df %>% dplyr::filter(weight > cutoff_include_all_ligands)

ligands_to_remove = setdiff(active_ligand_target_links_df$ligand %>% unique(), active_ligand_target_links_df_circos$ligand %>% unique())
targets_to_remove = setdiff(active_ligand_target_links_df$target %>% unique(), active_ligand_target_links_df_circos$target %>% unique())
  
circos_setup$circos_links_tmp = active_ligand_target_links_df %>% dplyr::filter(!target %in% targets_to_remove &!ligand %in% ligands_to_remove)

rm(active_ligand_target_links_df,active_ligand_target_links_df_circos,ligands_to_remove,targets_to_remove)


#### prepare the circos visualization


grid_col_ligand =c("Ligands" = "steelblue2")
grid_col_target =c("cDC2-DEGs" = "tomato")

def_circos_grid_col = function(circos_links,grid_col_ligand,grid_col_target) {
  grid_col_tbl_ligand = tibble(ligand_type = grid_col_ligand %>% names(), color_ligand_type = grid_col_ligand)
  grid_col_tbl_target = tibble(target_type = grid_col_target %>% names(), color_target_type = grid_col_target)
  # extra space: make a difference between a gene as ligand and a gene as target!
  circos_links = circos_links %>% mutate(ligand = paste0(ligand," "))
  circos_links = circos_links %>% inner_join(grid_col_tbl_ligand) %>% inner_join(grid_col_tbl_target)
  links_circle = circos_links %>% dplyr::select(ligand,target, weight)
  ligand_color = circos_links %>% distinct(ligand,color_ligand_type)
  grid_ligand_color = ligand_color$color_ligand_type %>% set_names(ligand_color$ligand)
  target_color = circos_links %>% distinct(target,color_target_type)
  grid_target_color = target_color$color_target_type %>% set_names(target_color$target)

  grid_col =c(grid_ligand_color,grid_target_color)
  return(list(grid_col = grid_col,
              circos_links = circos_links,
              links_circle = links_circle))
}

circos_setup$grid_col = def_circos_grid_col(circos_links = circos_setup$circos_links_tmp,
                                                    grid_col_ligand = grid_col_ligand,
                                                    grid_col_target = grid_col_target)

# give the option that links in the circos plot will be transparant ~ ligand-target potential score
circos_setup$transparency =  circos_setup$grid_col$circos_links %>% mutate(weight =(weight-min(weight))/(max(weight)-min(weight))) %>% mutate(transparency = 1-weight) %>% .$transparency


#### order ligands

target_order = circos_setup$grid_col$circos_links$target %>% unique()
ligand_order = circos_setup$grid_col$circos_links$ligand %>% intersect(circos_setup$grid_col$circos_links$ligand)
circos_setup$order = c(ligand_order,target_order)
rm(target_order,ligand_order)

#### define gaps

width_same_cell_same_ligand_type = 0.5
width_different_cell = 6
width_ligand_target = 15
width_same_cell_same_target_type = 0.5

circos_setup$gaps = c(
  # width_ligand_target,
  rep(width_same_cell_same_ligand_type, times = (circos_setup$grid_col$circos_links %>% dplyr::filter(ligand_type == "Ligands") %>% distinct(ligand) %>% nrow() -1)),
  width_ligand_target,
  rep(width_same_cell_same_target_type, times = (circos_setup$grid_col$circos_links %>% dplyr::filter(target_type == "cDC2-DEGs") %>% distinct(target) %>% nrow() -1)),
  width_ligand_target
  )

rm(width_same_cell_same_ligand_type, width_different_cell,width_ligand_target, width_same_cell_same_target_type)


#### render circos

library(circlize)
# circos.par(gap.degree = circos_setup$gaps)
# chordDiagram(circos_setup$grid_col$links_circle, 
#              directional = 1,
#              order=circos_setup$order,
#              link.sort = TRUE, 
#              link.decreasing = FALSE, 
#              grid.col = circos_setup$grid_col$grid_col,
#              transparency = 0, 
#              diffHeight = 0.005, 
#              direction.type = c("diffHeight", "arrows"),
#              link.arr.type = "big.arrow", 
#              link.visible = circos_setup$grid_col$links_circle$weight >= cutoff_include_all_ligands,
#              annotationTrack = "grid",
#              preAllocateTracks = list(track.height = 0.075))
# 
# # we go back to the first track and customize sector labels
# circos.track(track.index = 1, panel.fun = function(x, y) {
#     circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
#         facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.55), cex = 1)
# }, bg.border = NA)
# 
# circos.clear()
```

## circos plot and expression bar

```{r prepare gene expression}
target_cell_gene = list()
gene_color = data.frame()

for (type in c("Ligands","cDC2-DEGs")){

  if(type == "cDC2-DEGs") {
    target_gene = circos_setup$grid_col$circos_links$target %>% unique() %>% gsub(" ", "", ., fixed = TRUE)
    target_color = circos_setup$grid_col$circos_links$color_target_type[1]
  } else {
    target_gene = subset(circos_setup$grid_col$circos_links,ligand_type == type)
    target_color = target_gene$color_ligand_type[1]
    target_gene = target_gene$ligand %>% unique() %>% gsub(" ", "", ., fixed = TRUE)
  }

    if(type == "Ligands"){
    target_cell = subset(lung_scRNA_mouse_Raredon,subset = seurat_clusters == 9)
  } else{
    target_cell = subset(lung_scRNA_mouse_Raredon,subset = seurat_clusters == 5 & Gpr183.exp != 0)
  }

  target_cell_gene[[type]] = target_cell@assays$RNA@data %>% as.data.frame() %>% .[target_gene,]

  gene_mean = rowMeans(target_cell_gene[[type]]) %>% as.data.frame()
  gene_mean$type = type

  # ## correct ligand name accordingly
  gene_mean$ligand_name = ifelse(gene_mean$type != "cDC2-DEGs", paste0(rownames(gene_mean)," "), rownames(gene_mean))

  gene_color = rbind(gene_color,gene_mean)
}
```

# figures for the manuscript

## seurat figures

```{r plot figures for ms2}
## define sender cells
# pdf(file = paste0(working.dir,"MS_figures/Fig_JYu_11_Raredon_sender.pdf"),width = 6,height = 8)

DotPlot(lung_scRNA_mouse_Raredon, features = c("Pdgfra","Pi16","Col14a1", "Col1a1"))

# dev.off()

## define receiver cells
# pdf(file = paste0(working.dir,"MS_figures/Fig_JYu_12_Raredon_receiver.pdf"),width = 6,height = 8)

DotPlot(lung_scRNA_mouse_Raredon, features = c("Itgae", "Cd86","Mgl2","Gpr183", "Ccl17"))

# dev.off()

## umap of the seurat object
## highlight cluster 5 and 9
# pdf(file = paste0(working.dir,"/MS_figures/Fig_JYu_14_Raredon_umap.pdf"))
DimPlot(lung_scRNA_mouse_Raredon,label = TRUE,group.by = "seurat_clusters",cols = c(rep("grey",5),"tomato",rep("grey",3),"steelblue2",rep("grey",18)))
# dev.off()

# pdf(file = paste0(working.dir,"MS_figures/Fig_JYu_16_Raredon_Tslp_Crlf2.pdf"))
DotPlot(lung_scRNA_mouse_Raredon,features = c("Tslp","Crlf2"))
# dev.off()
```

## heatmap: nichenetr ligand-receptor 

```{r plot figures for ms3}
## heatmap of ligand and receptors

# pdf(file = paste0(working.dir,"MS_figures/Fig_JYu_13_Raredon_ligand_receptor.pdf"),width = 10,height = 6)

Raredon.fibro_Gpr183cDC2_all$p_ligand_receptor_network

# dev.off()
```


## circos plot

```{r plot figures for ms4,message=FALSE,warning=FALSE}
# pdf(file = paste0(working.dir,"MS_figures/Fig_JYu_15_Raredon_circosplot.pdf"),width = 15,height = 15)

circos.par(gap.degree = circos_setup$gaps,track.margin=c(0,0),canvas.xlim=c(-1.5,1.5),canvas.ylim=c(-1.5,1.5))
chordDiagram(circos_setup$grid_col$links_circle, 
             directional = 1,
             order=circos_setup$order,
             link.sort = TRUE, 
             link.decreasing = FALSE, 
             grid.col = circos_setup$grid_col$grid_col,
             # col = circos_setup$grid_col$grid_col,
             transparency = 0, 
             diffHeight = 0.005, 
             direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", 
             link.visible = circos_setup$grid_col$links_circle$weight >= cutoff_include_all_ligands,
             annotationTrack = "grid",
             preAllocateTracks = list(track.height = 0.075))

## add extra information and annotation to the outer track 
circos.track(track.index = 1,
             ylim = c(-1, 1),
             bg.border = NA,
             panel.fun = function(x, y) {
               value = subset(gene_color, ligand_name == CELL_META$sector.index)
               
               ## add barplot for gene expression
               circos.barplot(value = value[,1], pos = CELL_META$xcenter, col = "grey")
               
               ## add gene name
               circos.text(x=CELL_META$xcenter,y=CELL_META$ylim[1]+5,labels=CELL_META$sector.index,facing = "clockwise",niceFacing = TRUE, adj = c(0, 0.4), cex = 1)
  }
)

circos.clear()

# dev.off()
```

